Predicting instabilities of a tuneable ring laser with an iterative map model

### 2020-11-22 Papers/Journal/Tuneable_Lasers.tex
Updated references

\section{Introduction}
\label{sec:intro}
Tuneable lasers have the ability to vary the frequency of their output by up to about 100 nanometres~\cite{bohun2015, burgoyne2010, yamashita2009}. This ability to vary the frequency leads to applications such as optical coherence tomography~\cite{bohun2015, burgoyne2014, yamashita2009}, coherent anti-Stokes Raman spectroscopy~\cite{burgoyne2014}, deep tissue multi-photon microscopy~\cite{chung2017, liu2017}, and diagnostics of ultrafast processes~\cite{burgoyne2014, silfvast2004}. Tuneable lasers are usually constructed in a ring with five main components: output coupler, chirped fibre Bragg grating (CFBG), modulator, Er-doped fibre, and pump laser. A typical ring laser cavity is depicted in Fig.~\ref{fig:cavity}. Due to the high power and short duration of the pulses, the Kerr nonlinearity plays a crucial role in the dynamics within the cavity.

Several effects arise within the cavity due to the interplay of dispersion, modulation, and the nonlinearity~\cite{bohun2015, coen1997, lapre2019, shao2019, woodward2018}. The two effects of most interest for this paper are wave breaking, and modulation instability. Wave breaking causes the leading edge of a pulse to be red shifted, and the trailing edge blue shifted (in the anomalous dispersion regime, the effect is reversed) to the point that a shock is developed~\cite{anderson1992, rothenberg1989a, rothenberg1989b, tomlinson1984, tomlinson1985}. The shifts in frequencies cause the pulse to become more rectangular in the frequency domain with a linear chirp over most of the pulse. In optics, wave breaking manifests from self-phase modulation (SPM) which is a direct effect of the Kerr nonlinearity causing the pulse to interfere with itself. New frequencies are then generated which induces the rectangular profile~\cite{agrawal2013, mollenauer1980, woodward2018}. Outside of optics, wave breaking occurs in other areas with nonlinear waves such as plasmas, heat propagation, transmission lines, and fluid dynamics~\cite{rothenberg1989b}. Modulation instability---which typically arises in the anomalous dispersion regime---is the other effect that will be of interest. However, in the presence of multiple frequencies, modulation instability can emerge in the normal dispersion regime as well through cross-phase modulation and four-wave mixing~\cite{agrawal1987, agrawal1989, agrawal2013, haelterman1992}. Modulation instability causes a wave to break into short pulses, and can cause a laser to lose coherence~\cite{agrawal1987, coen1997, haelterman1992}. Within a ring laser cavity wave breaking and modulation instability can become parasitic leading to an unstable and unsustainable pulse. This gives rise to the need of understanding the rich landscape of the parameter space, and determining design principles to guarantee the ring laser is stable and sustainable~\cite{bohun2015, burgoyneemail, finot2008, lapre2019, woodward2018}.

\begin{figure}[tbp]
	\centering
	\input{Cavity}
	\caption{Typical cavity of a fibre ring laser~\cite{burgoyne2014, chung2017, lapre2019, shao2019, tang2014}. The laser pulses travel clockwise around each loop.}
	\label{fig:cavity}
\end{figure}

@@

% Table of parameter values
 \begin{table*}[tbp]
	\centering
	\caption{Range of variation of various parameters.}
 	\label{tab:values}
 	\begin{tabular}{lcll}
 		\hline\noalign{\smallskip}
 		Parameter & Symbol & Value & Sources \\
 		\hline\noalign{\smallskip}
 		Absorption of Fibre & $\alpha$ & $10^{-4}$--$0.3\text{ m}^{-1}$  &~\cite{burgoyneemail, shtyrina2017, tomlinson1984, usechak2005, yarutkina2013} \\
 		Fibre Dispersion & $\beta_2^f$ & $-50$--$50 \text{ ps}^2/ \text{km}$ &~\cite{agrawal2002, agrawal2013, burgoyne2014, litchinitser1997, peng2018, yarutkina2013} \\
 		Fibre Nonlinearity & $\gamma$ & $0.001$--$0.01 \text{ W}^{-1} \text{m}^{-1}$ &~\cite{agrawal2013, finot2008, usechak2005, yarutkina2013} \\
 		Grating Dispersion & $\beta_2^g L_D$ & $10$--$2000 \text{ ps}^2$ &~\cite{agrawal2002, agrawal2013, burgoyne2014, li1998} \\
 		Length of Cavity & $L_T$ & $10$--$100 \text{ m}$ &~\cite{burgoyneemail, peng2018, tamura1996} \\
 		Length of Fibre & $L_f$ & $0.15$--$1 \text{ m}$ &~\cite{burgoyneemail} \\
 		Length of Gain Fibre & $L_g$ & $2$--$3 \text{ m}$ &~\cite{burgoyne2014, peng2018, shtyrina2017, tamura1993,yarutkina2013} \\
 		Modulation Time & $T_M$ & $15$--$150 \text{ ps}$ &~\cite{bohun2015, burgoyneemail, burgoyne2014} \\
 		Reflectivity of Optical Coupler & $R$ & $0.1$--$0.9$ &~\cite{burgoyneemail, li1998, peng2018,  tamura1993, tamura1996, yamashita2009} \\
 		Saturation Energy & $\Es$ & $10^3$--$10^4 \text{ pJ}$ &~\cite{burgoyneemail, usechak2005, yarutkina2013} \\
 		Small Signal Gain & $g_0$ & $1$--$10 \text{ m}^{-1}$ &~\cite{burgoyneemail, yarutkina2013} \\
 		\noalign{\smallskip}\hline
 	\end{tabular}
 \end{table*}

@@

\subsection{Nonlinear Solution and Instability}
\label{sec:nlresults}
In this iterative scheme---as well as within the laboratory---we must specify the input pulse shape. One of the most common form is a hyperbolic secant~\cite{coen1997, finot2008, mitschke1986, rothenberg1989b, tomlinson1984}, which we assume has the exact form
\begin{align}
	A_0 = \Gamma \sech \left( 2 T \right) \textrm{e}^{i \pi / 4},
	\label{eq:nlA0}
\end{align}
where $\Gamma$ is a normalizing factor chosen so the pulse has the initial energy $E_0$, which we take as $E_0 = 0.1$, unless specified otherwise.

>>>>

\section{Introduction}
\label{sec:intro}
Tuneable lasers have the ability to vary the frequency of their output by up to about 100 nanometres~\cite{bohun2015, burgoyne2010, yamashita2009}. This ability to vary the frequency leads to applications such as optical coherence tomography~\cite{bohun2015, burgoyne2014, yamashita2009}, coherent anti-Stokes Raman spectroscopy~\cite{burgoyne2014}, deep tissue multi-photon microscopy~\cite{chung2017}, and diagnostics of ultrafast processes~\cite{burgoyne2014, silfvast2004}. Tuneable lasers are usually constructed in a ring with five main components: output coupler, chirped fibre Bragg grating (CFBG), modulator, Er-doped fibre, and pump laser. A typical ring laser cavity is depicted in Fig.~\ref{fig:cavity}. Due to the high power and short duration of the pulses, the Kerr nonlinearity plays a crucial role in the dynamics within the cavity.

Several effects arise within the cavity due to the interplay of dispersion, modulation, and the nonlinearity~\cite{bohun2015, coen1997, lapre2019, shao2019, woodward2018}. The two effects of most interest for this paper are wave breaking, and modulation instability. Wave breaking causes the leading edge of a pulse to be red shifted, and the trailing edge blue shifted (in the anomalous dispersion regime, the effect is reversed) to the point that a shock is developed~\cite{anderson1992, rothenberg1989a, rothenberg1989b, tomlinson1984, tomlinson1985}. The shifts in frequencies cause the pulse to become more rectangular in the frequency domain with a linear chirp over most of the pulse. In optics, wave breaking manifests from self-phase modulation (SPM) which is a direct effect of the Kerr nonlinearity causing the pulse to interfere with itself. New frequencies are then generated which induces the rectangular profile~\cite{agrawal2013, woodward2018}. Outside of optics, wave breaking occurs in other areas with nonlinear waves such as plasmas, heat propagation, transmission lines, and fluid dynamics~\cite{rothenberg1989b}. Modulation instability---which typically arises in the anomalous dispersion regime---is the other effect that will be of interest. However, in the presence of multiple frequencies, modulation instability can emerge in the normal dispersion regime as well through cross-phase modulation and four-wave mixing~\cite{agrawal1987, agrawal2013, haelterman1992}. Modulation instability causes a wave to break into short pulses, and can cause a laser to lose coherence~\cite{agrawal1987, coen1997, haelterman1992}. Within a ring laser cavity wave breaking and modulation instability can become parasitic leading to an unstable and unsustainable pulse. This gives rise to the need of understanding the rich landscape of the parameter space, and determining design principles to guarantee the ring laser is stable and sustainable~\cite{bohun2015, burgoyneemail, finot2008, lapre2019, woodward2018}.

\begin{figure}[tbp]
	\centering
	\input{Cavity}
	\caption{Typical cavity of a fibre ring laser~\cite{burgoyne2014, chung2017, lapre2019, shao2019}. The laser pulses travel clockwise around each loop.}
	\label{fig:cavity}
\end{figure}

@@

% Table of parameter values
\begin{table*}[tbp]
	\centering
	\caption{Range of variation of various parameters.}
 	\label{tab:values}
 	\begin{tabular}{lcll}
		\hline\noalign{\smallskip}
		Parameter & Symbol & Value & Sources \\
		\hline\noalign{\smallskip}
		Absorption of Fibre & $\alpha$ & $10^{-4}$--$0.3\text{ m}^{-1}$  &~\cite{burgoyneemail, shtyrina2017, tomlinson1984, usechak2005} \\
		Fibre Dispersion & $\beta_2^f$ & $-50$--$50 \text{ ps}^2/ \text{km}$ &~\cite{agrawal2002, agrawal2013, burgoyne2014} \\
		Fibre Nonlinearity & $\gamma$ & $0.001$--$0.01 \text{ W}^{-1} \text{m}^{-1}$ &~\cite{agrawal2013, finot2008, usechak2005} \\
		Grating Dispersion & $\beta_2^g L_D$ & $10$--$2000 \text{ ps}^2$ &~\cite{agrawal2002, agrawal2013, burgoyne2014} \\
		Length of Cavity & $L_T$ & $10$--$100 \text{ m}$ &~\cite{burgoyneemail, tamura1996} \\
		Length of Fibre & $L_f$ & $0.15$--$1 \text{ m}$ &~\cite{burgoyneemail} \\
		Length of Gain Fibre & $L_g$ & $2$--$3 \text{ m}$ &~\cite{burgoyne2014, shtyrina2017} \\
		Modulation Time & $T_M$ & $15$--$150 \text{ ps}$ &~\cite{bohun2015, burgoyneemail, burgoyne2014} \\
		Reflectivity of Optical Coupler & $R$ & $0.1$--$0.9$ &~\cite{burgoyneemail, tamura1996, yamashita2009} \\
		Saturation Energy & $\Es$ & $10^3$--$10^4 \text{ pJ}$ &~\cite{burgoyneemail, usechak2005} \\
		Small Signal Gain & $g_0$ & $1$--$10 \text{ m}^{-1}$ &~\cite{burgoyneemail} \\
		\noalign{\smallskip}\hline
	\end{tabular}
\end{table*}

@@

\subsection{Nonlinear Solution and Instability}
\label{sec:nlresults}
In this iterative scheme---as well as within the laboratory---we must specify the input pulse shape. One of the most common form is a hyperbolic secant~\cite{coen1997, finot2008, rothenberg1989b, tomlinson1984}, which we assume has the exact form
\begin{align}
	A_0 = \Gamma \sech \left( 2 T \right) \textrm{e}^{i \pi / 4},
	\label{eq:nlA0}
\end{align}
where $\Gamma$ is a normalizing factor chosen so the pulse has the initial energy $E_0$, which we take as $E_0 = 0.1$, unless specified otherwise.
